---
title: "Maryland's Volleyball Points Explained"
author: "Mauriah Allen"
date: '2024-10-11'
format:
  html:
    code-fold: true
---

###Load Libraries

```{r}
library(tidyverse)
library(waffle)

matches24 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/main/data/ncaa_womens_volleyball_matchstats_2024.csv")

```

### Seperating Maryland's Data

```{r}
md_data <- matches24 |> filter(team == "Maryland")
```

### Calculating the total number of kills, blocks, aces and opponent errors for Maryland

```{r}

total_kills <- sum(md_data$kills)
total_blocks <- sum(md_data$block_solos)
total_aces <- sum(md_data$aces)
total_opponent_errors <- sum(md_data$defensive_errors)

md <- c("total_kill"=716, "total_blocks"=25, "total_aces"=127, "total_opponent_error"=322)
```

### Making a waffle chart for Maryland kills,blocks,aces and opponent errors.

```{r}
waffle(
        md/2, 
        rows = 10, 
        title="Maryland's stats", 
        xlab="1 square = 1 yard", 
        colors = c("green", "red","orange", "blue")
)
```

### Creating a waffle chart for Nebraska to Compare

```{r}
nb_data <- matches24 |> filter(team == "Nebraska")

total_kills <- sum(nb_data$kills)
total_blocks <- sum(nb_data$block_solos)
total_aces <- sum(nb_data$aces)
total_opponent_errors <- sum(nb_data$defensive_errors)

nb <- c("total_kill"=701, "total_blocks"=23, "total_aces"=71, "total_opponent_error"=283)

waffle(
        nb/2, 
        rows = 10, 
        title="Nebraska's stats", 
        xlab="1 square = 1 yard", 
        colors = c("green", "red","orange", "blue")
)
```

Comparing the differences between Maryland and Nebraska

```{r}
md <- c("total_kill"=716, "total_blocks"=25, "total_aces"=127, "total_opponent_error"=322)
nb <- c("total_kill"=701, "total_blocks"=23, "total_aces"=71, "total_opponent_error"=283, 112)

```

```{r}
waffle(
        nb/2, 
        rows = 10, 
        title="Nebraska's stats", 
        xlab="1 square = 1 yard", 
        colors = c("green", "red","orange", "blue", "white")
)

waffle(
        md/2, 
        rows = 10, 
        title="Maryland's stats", 
        xlab="1 square = 1 yard", 
        colors = c("green", "red","orange", "blue")
)
```

In conclusion, most of Maryland's points go towards kills, as this is the case for many other schools. Compared to Nebraska, Maryland has 15 more kills than Nebraska.They have around the same number of blocks, which showed me that blocks aren't a regular statistic to gain. Aces though, Maryland is leading Nebraska with a 56 point deference. And in opponent errors it looks like Maryland's opponents have made 39 more mistakes than Nebraska's opponents. Maryland really thrives off their kills and aces.
